Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-25 WSSSPE3/report/wg_appendix_eng_design.tex
Fix typos.

The group included members from different backgrounds, including quantum chemistry, epidemiology, computer science, software engineering, and microscopy. Each participant was invited to give their perspective on the topic area and what they thought were the crucial points for discussion. There was a general consensus that there was a need for relating principles to practice for the computational science and engineering community. Furthermore, various members of the group expressed their interest in tools and best practices for facilitating the maintenance and evolution of scientific software systems. It was agreed to identify principles from software engineering and from sustainability design and, based on those lists, discuss what each of those would mean applied to specific example systems from the expert domains of some of the group members. The group identified a number of software engineering principles drawn from the SoftWare Engineering Body of Knowledge (SWEBOK)~\cite{swebokv3}.

@@

\subsection{Description of Funding Needed}
Specific funding was not discussed in this working group. However, this is a open topic that can be be discussed in relation to emerging funding calls from National agencies or grant proposal initiatives.

>>>>

The group included members from different backgrounds, including quantum chemistry, epidemiology, computer science, software engineering, and microscopy. Each participant was invited to give their perspective on the topic area and what they thought were the crucial points for discussion. There was a general consensus that there was a need for relating principles to practice for the computational science and engineering community. Furthermore, various members of the group expressed their interest in tools and best practices for facilitating the maintenance and evolution of scientific software systems. It was agreed to identify principles from software engineering and from sustainability design and, based on those lists, discuss what each of those would mean applied to specific example systems from the expert domains of some of the group members. The group identified a number of software engineering principles drawn from the Software Engineering Body of Knowledge (SWEBOK)~\cite{swebokv3}.

@@

\subsection{Description of Funding Needed}
Specific funding was not discussed in this working group. However, this is a
open topic that can be discussed in relation to emerging funding calls from
National agencies or grant proposal initiatives.

### 2016-01-25 WSSSPE3/report/wg_appendix_funding_spec_expert.tex
Fix typos.

One model that has been helpful in finding funding for RSE groups is the use of
funds left over on research grants when RAs have left prematurely -- PIs like
this arrangement as it is hard to find good staff for short-term positions, so
having a pool of research programming staff on hand resolves this problem. We
recommend that funders give explicit guidance to grant holders and institutions
that such arrangement are favorable. Framework agreements permitting this to go
ahead without checking back every time with funders and/or grant panels would
further smooth this. (This also provides more stable jobs for those who hold
these skills, but arguments about making life nicer for postdocs will not help
persuade funders or PIs!)

There is some question about the most effective duration and percent of full
time for a programmer's work on a project. At least three months is necessary
for the programmer to read into the science (RSEs must not become so disengaged
from research that they do not have time to read a few papers -- this will result
in code which does not meet scientific needs), but too long could result in an
RSE losing their flexibility, becoming so engaged in one project that when that
project ends, they find it hard to transfer. For this reason, we also recommend
that 40 percent is ideal; two projects per developer, with some time for
training and infrastructure work. Two developers per project seems to be ideal,
in the sense that software development is enhanced by two pairs of eyes.

>>>>

One model that has been helpful in finding funding for RSE groups is the use of
funds left over on research grants when RAs have left prematurely -- PIs like
this arrangement as it is hard to find good staff for short-term positions, so
having a pool of research programming staff on hand resolves this problem. We
recommend that funders give explicit guidance to grant holders and institutions
that such arrangement is favorable. Framework agreements permitting this to go
ahead without checking back every time with funders and/or grant panels would
further smooth this. (This also provides more stable jobs for those who hold
these skills, but arguments about making life nicer for postdocs will not help
persuade funders or PIs!)

There is some question about the most effective duration and percent of full
time for a programmer's work on a project. At least three months is necessary
for the programmer to read into the science (RSEs must not become so disengaged
from research that they do not have time to read a few papers -- this will result
in code which does not meet scientific needs), but too long could result in an
RSE losing their flexibility, becoming so engaged in one project that when that
project ends, they find it hard to transfer. For this reason, we also recommend
that 40 percent is ideal; two projects per developer, with some time for
training and infrastructure work. Having two developers per project seems to be ideal,
in the sense that software development is enhanced by two pairs of eyes.
